---
title: "Revealing Alien Patterns: A Data-Driven Investigation into Global UFO Sightings"
subtitle: "Decoding the Mysteries: Analyzing Temporal, Spatial, and Descriptive Patterns in Over 80,000 UFO Sightings"
format: 
  html:
    embed-resources: true
toc: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

# Project Overview

The fascination with unidentified flying objects (UFOs) has captured the imagination of humanity for decades. This project aims to explore a comprehensive dataset comprising over 80,000 records of UFO sightings dating back to 1949. By leveraging the wealth of information provided, including latitude and longitude data, date and time stamps, and detailed descriptions of each sighting, we seek to unravel patterns, trends, and insights regarding UFO activities around the world.

# Dataset Overview

**Reasons for Choosing the Dataset:**

The dataset under consideration comes originally from [NUFORC](http://www.nuforc.org/) contains over 80,000 records of UFO sightings dating back to 1949, making it a rich source for exploring the temporal and spatial patterns of extraterrestrial activities. The inclusion of latitude and longitude data, date and time stamps, and detailed descriptions of each sighting provides a comprehensive foundation for conducting a thorough analysis.

```{r}

# Load necessary libraries
library(tidyverse)

ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv", show_col_types = FALSE)
```

```{r}

# Display basic information about the dataset
#cat("Dataset Information:\n")
cat("========== Dataset Structure ==========\n")
str(ufo_sightings)

# Display summary statistics
#cat("\nSummary Statistics:\n")
#summary(ufo_sightings) --- Dropped Summary upon receiving feedback

#Installing and Loading the dlookr package

library(dlookr)

# Showing the aspects of the columns of dataset using dlookr package
overview <- diagnose(ufo_sightings)

# Print the overview
cat("========== Dataset Overview ==========\n")

print(overview)

# Display the dimensions of the dataset
#cat("\nDimensions of the Dataset:\n")
cat("========== Dataset Dimensions ==========\n")
#dim(ufo_sightings)
cat("Number of rows:", nrow(ufo_sightings), "\n")
cat("Number of columns:", ncol(ufo_sightings), "\n")

# Display the first few rows of the dataset
#cat("\nFirst Few Rows of the Dataset:\n")
cat("========== First Few Rows of the Dataset ==========\n")
head(ufo_sightings)
```

## Questions 

**Question 1: Global Distribution Analysis**
**Question 2: Temporal Trends by Time of Day**

## Analysis plan
A plan for answering each of the questions including the variables involved, variables to be created , external data to be merged in (if any).


**Variables Involved to answer Question1:** 
latitude- The latitude for this city, from geonames.org
longitude- The longitude for this city, from geonames.org
state-The state, province, or similar division of the sighting.

No new variables are required to answer question1.

** Plan for answering question1 **
1.Data Cleaning -Ensure latitude, longitude and state columns are clean and contain no missing or outlier values.
2.Visualization - Creating a global map of UFO Sightings using ggplot.
3.Identify hotspot- Analyze the map to observe geographical hostpots with high UFO activity and ranking the regions with high UFO Sightings using descriptive statistics.
4.Regional Trends - Explore whether there are specific patterns or trends associated with certain states or countries.

**Variables Involved to answer Question2:**
Time of Day -New variable for time of day binning to be build using date_time variable.

**Plan for answering question2 **
1.Data Preparation- Converting the existing data and time stamps into format suitable for analysis and create new variable that indicates time of day.
2.Visualization - Creating a time series plot to visualize the overall trend of UFO sightings over time and generating bar graph to represent distribution of sightings across different times of day.
3.Temporal patterns- Conduct statistical analysis to identify if there are significant differences in sighting frequencies during specific times of the day.
4.Interpretation- Consider if there are correlations between specific times of the day and the frequency or characteristics of sightings.

# Expected Outcomes

1.  **Spatial Insights:**

    -   Visualization of global UFO sighting distribution.

    -   Identification of regions with high UFO activity.

2.  **Temporal Patterns:**

    -   Uncovering patterns in the timing and duration of UFO sightings.

# Conclusion

This project seeks to bring a data-driven approach to the mysterious realm of UFO sightings. By analyzing historical data, we aim to uncover patterns, answer questions about temporal and spatial characteristics, and contribute to our understanding of these extraterrestrial phenomena. The findings may have public awareness, providing valuable insights into the unknown.

## References

1. Title: UFO Sightings Redux
   Author: [jonthegeek](https://github.com/jonthegeek)
   Date: 2023-06-20
   Link: "https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-06-20"
