---
title: "Revealing Alien Patterns: A Data-Driven Investigation into Global UFO Sightings"
subtitle: "Decoding the Mysteries: Analyzing Temporal, Spatial, and Descriptive Patterns in Over 80,000 UFO Sightings"
format: 
  html:
    embed-resources: true
toc: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

# Project Overview

The fascination with unidentified flying objects (UFOs) has captured the imagination of humanity for decades. This project aims to explore a comprehensive dataset comprising over 80,000 records of UFO sightings dating back to 1949. By leveraging the wealth of information provided, including latitude and longitude data, date and time stamps, and detailed descriptions of each sighting, we seek to unravel patterns, trends, and insights regarding UFO activities around the world.

# Dataset Overview

**Reasons for Choosing the Dataset:**

The dataset under consideration comes originally from [NUFORC](http://www.nuforc.org/) contains over 80,000 records of UFO sightings dating back to 1949, making it a rich source for exploring the temporal and spatial patterns of extraterrestrial activities. The inclusion of latitude and longitude data, date and time stamps, and detailed descriptions of each sighting provides a comprehensive foundation for conducting a thorough analysis.

```{r}

# Load necessary libraries
library(tidyverse)

ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")
```

```{r}

# Display basic information about the dataset
cat("Dataset Information:\n")
str(ufo_sightings)

# Display summary statistics
cat("\nSummary Statistics:\n")
summary(ufo_sightings)

# Display the dimensions of the dataset
cat("\nDimensions of the Dataset:\n")
dim(ufo_sightings)

# Display the first few rows of the dataset
cat("\nFirst Few Rows of the Dataset:\n")
head(ufo_sightings)
```

# Questions to Answer

**Question 1: Global Distribution Analysis - where are the UFO sightings?**

**Plan:** Utilize latitude, longitude & State data to visualize the global distribution of UFO sightings. Identify geographical hotspots and trends to understand if there are regions more prone to UFO activities.

**Variables Involved:** Latitude, Longitude & State

**Question 2: Temporal Trends by Time of Day - when do we have UFO sightings?**

**Plan:** Examine if there are specific times of the day when UFO sightings are more prevalent. Utilize the time data from the dataset. Create time bins (e.g., morning, afternoon, evening, night) and analyze sighting patterns within each bin. Investigate if there are temporal trends associated with different times of the day.

**Variables Involved:** Time of Day (New variable for time of day binning to be build using date_time variable).

# Expected Outcomes

1.  **Spatial Insights:**

    -   Visualization of global UFO sighting distribution.

    -   Identification of regions with high UFO activity.

2.  **Temporal Patterns:**

    -   Uncovering patterns in the timing and duration of UFO sightings.

# Conclusion

This project seeks to bring a data-driven approach to the mysterious realm of UFO sightings. By analyzing historical data, we aim to uncover patterns, answer questions about temporal and spatial characteristics, and contribute to our understanding of these extraterrestrial phenomena. The findings may have public awareness, providing valuable insights into the unknown.
